---
title: "HCV predection with K-means Clustering"
Auther: Waleed ALjomah
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r }
# Helper packages
library(dplyr)       # for data manipulation
library(ggplot2)     # for data visualization
library(stringr)     # for string functionality

# Modeling packages
library(cluster)     # for general clustering algorithms
library(factoextra)  # for visualizing cluster results
```



###data : http://archive.ics.uci.edu/ml/datasets/HCV+data
###publication : https://www.researchgate.net/profile/Georg_Hoffmann4/publication/326151919_Using_machine_learning_techniques_to_generate_laboratory_diagnostic_pathways-a_case_study/links/5b3e3e5b0f7e9b0df5f58d31/Using-machine-learning-techniques-to-generate-laboratory-diagnostic-pathways-a-case-study.pdf

The data set contains laboratory values of blood donors and Hepatitis C patients
and demographic values like age.

```{r }
#import data
HCV <- read.csv("data/hcvdat0.csv")
```

This data set is a Multivariate with supervised learning with Classification and clustering problem

```{r}
#Explore data

head(HCV)

summary(HCV)
```

All attributes except Category and Sex are numerical. The laboratory data are the attributes 5-14.
1) X (Patient ID/No.)
2) Category (diagnosis) (values: '0=Blood Donor', '0s=suspect Blood Donor', '1=Hepatitis', '2=Fibrosis', '3=Cirrhosis')
3) Age (in years)
4) Sex (f,m)
5) ALB
6) ALP
7) ALT
8) AST
9) BIL
10) CHE
11) CHOL
12) CREA
13) GGT
14) PROT

##Feature engeneering :

- there is a missing values which should be deleted for the purpose of modeling

##distribution of missing values in a data
```{r}
#visualising missing

HCV %>%
 is.na() %>%
 reshape2::melt() %>%
 ggplot(aes(Var2, Var1, fill=value)) +
 geom_raster() +
 coord_flip() +
 scale_y_continuous(NULL, expand = c(0, 0)) +
 scale_fill_grey(name = "",
                 labels = c("Present",
                            "Missing")) +
 xlab("Observation") +
 theme(axis.text.y  = element_text(size = 4))
```

- missing values mainly in "ALP" and "CHOL" 


```{r}
#dealing with missing

HepC <- na.omit(HCV)
HepC
```

### to convert any ordinal categorical variables to numeric and one-hot encode the remaining nominal categorical variables.
```{r}


# One-hot encode --> retain only the features and not "Category"
full_rank  <- caret::dummyVars(Category ~ ., data = HepC, 
                               fullRank = TRUE)
HepC_1hot <- predict(full_rank, HepC)

# Scale data
HepC_1hot_scaled <- scale(HepC_1hot)

# New dimensions
dim(HepC_1hot_scaled)

```
### determining how many cluster to apply by using  "elbow method"
by computing k-means clustering for different values of k, then plotting the curve of WSS and determing the location of a bend (i.e. elbow) which is an indicator of the appropriate number of clusters.

```{r}
set.seed(123)

HepC_clustering <- fviz_nbclust(
  HepC_1hot_scaled, 
  kmeans, 
  method = "wss", #within-cluster sum of squares
  k.max = 25, 
  verbose = FALSE
)

HepC_clustering
```


number of clustering 12 or 13 may be the best fit for this model.






